\sysfull

### 2018-04-10 paper/intro.tex
last comments from OOPSLA 2018 branch by Alexei ported manually to main branch

Energy-harvesting devices create unique challenges because they operate {\em intermittently} when energy is available~\citep{hicks_isca_2017,lucia_snapl_2017}. An
energy-harvesting device buffers energy in a small storage, e.g. a capacitor~\citep{gorlatova_tmc_2013},~\citep{gunduz_commag_2014} and when a threshold amount of energy accumulates, begins operating. Harvestable energy sources are low-power compared to a platform's operating level. A device operates briefly and when buffered energy is depleted, shuts down and recharges to operate again later. As an example, recharge time may be tens of seconds in radio frequency powered medical device~\cite[Fig. 3c]{nadeau_naturebio_2017}. Moreover, charge and discharge times vary by device (due to different capacitor sizes) and some may fail $\approx$10 to $\approx$100 times per second~\citep{tan_infocom_2016},~\citep{mementos},~\citep{nvp}.

@@

There are two main approaches for dealing with data inconsistency for intermittently-powered devices: (i) \emph{programming and execution models}~\citep{dino,ratchet,chain,alpaca} and (ii) \emph{architectures}~\citep{hicks_isca_2017,idetic,nvp,tictpl}. New architectures require hardware changes and are inapplicable to today's systems~\citep{hicks_isca_2017,nvp}, therefore new programming models and compilers are more suitable (as of now). However, they have their own limitations that need to be addressed---the core one being the rigidity of the \emph{static} decomposition into tasks.

@@

\textbf{Task Decomposition of Intermittent Programs.} Recently proposed {\em
task-based} programming and execution models~\citep{alpaca,chain} advocate for
static decomposition of a program (by a programmer) into a collection of tasks.
Tasks can include arbitrary computation and, upon completion, are guaranteed to
have executed {\em atomically}, despite arbitrarily-timed power failures.
The programmer explicitly expresses task-to-task control flow. The execution
in Figure~\ref{fig:coalesce} (top) illustrates how the decomposition into
tasks affects the time the application takes to complete. Each task transition introduces the overhead of tracking and atomically committing the modifications to non-volatile memory, to maintain consistency of program state~\citep{chain,alpaca}. Naively minimizing the number of task transitions, (i.e. the task count), \emph{at compile time} risks creating a task that requires more energy than the device can buffer in its energy storage. Such a large task would fail to complete unless the environment provides sufficient incoming energy to supplement the energy in the capacitor. To eliminate this risk, the programmer (or the compiler) must decompose the program conservatively into many small tasks, at the cost of \emph{higher task transition overhead}.

One approach to decrease the task count, without assuming a minimum level of
incoming power always available, is to predict task energy consumption and
choose a decomposition specific to the device capacitor size. However, statically predicting energy of arbitrary input-dependent code with peripheral access is a problem without a general solution. Furthermore, a static decomposition approach prevents portability across devices with different storage capacitors.

@@

Task coalescing removes the burden from the programmer, because it accepts any task decomposition and improves it dynamically. To further reduce the programmer effort, we propose a compiler pass for \emph{automatic decomposition} of programs into (small) \emph{atomic} tasks. The compiler identifies non-volatile variables shared across tasks as tasks are created, and instruments reads and writes of those variables, using memory virtualization, to keep the data consistent in the presence of power loss. Despite being limited to a subset of the C language, the automatic task decomposition allowed us to port several
applications to an intermittent platform with a moderate effort.

@@

\begin{itemize}
	\item Two adaptive coalescing strategies for selecting boundaries to coalesce based on past execution behaviour;
	\item Software memory virtualisation for transiently-powered devices that preserves atomicity of coalesced tasks;
	\item Task splitting startegy that preserve forward progress and memory consistency;
\end{itemize}

>>>>

Energy-harvesting devices create unique challenges because they operate {\em intermittently} when energy is available~\citep{hicks_isca_2017,lucia_snapl_2017}. An
energy-harvesting device buffers energy in a small storage, e.g. a capacitor~\citep{gorlatova_tmc_2013},~\citep{gunduz_commag_2014} and when a threshold amount of energy accumulates, begins operating. Harvestable energy sources are low-power compared to a platform's operating level. A device operates briefly and when buffered energy is depleted, shuts down and recharges to operate again later. As an example, recharge time may be tens of seconds in radio frequency powered medical device~\cite[Fig. 3c]{nadeau_naturebio_2017}. Moreover, charge and discharge times vary by device (due to different capacitor sizes) and some may fail $\approx$10 to $\approx$100 times per second~\citep{tan_infocom_2016,mementos,nvp}.

@@

There are two main approaches for dealing with data inconsistency for intermittently-powered devices: (i) \emph{programming and execution models}~\citep{dino,ratchet,chain,alpaca} and (ii) \emph{architectures}~\citep{hicks_isca_2017,idetic,nvp,tictpl}. \textcolor{red}{tictpl is not HW, is it? also it doesn't deal with inconsistency; this paragraph doesn't include Hibernus, so it should not include CPTL either.} New architectures require hardware changes and are inapplicable to today's systems~\citep{hicks_isca_2017,nvp}, therefore new programming models and compilers are more suitable (as of now). However, they have their own limitations that need to be addressed---the core one being the rigidity of the \emph{static} decomposition into tasks.

@@

\textbf{Task Decomposition of Intermittent Programs.} Recently proposed {\em
task-based} programming and execution models~\citep{alpaca,chain} advocate for
static decomposition of a program (by a programmer) into a collection of tasks. \textcolor{red}{add \cite{intel} work; separate by-programmer and by-compiler; both cases hit the energy estimation problem; also intel work proposes a method for decomposing into one of many decompositions, but does not give a good solution for choosing one}
Tasks can include arbitrary computation and, upon completion, are guaranteed to
have executed {\em atomically}, despite arbitrarily-timed power failures.
The programmer explicitly expresses task-to-task control flow. The execution
in Figure~\ref{fig:coalesce} (top) illustrates how the decomposition into
tasks affects the time the application takes to complete. Each task transition introduces the overhead of tracking and atomically committing the modifications to non-volatile memory, to maintain consistency of program state~\citep{chain,alpaca}. Naively minimizing the number of task transitions, (i.e. the task count), \emph{at compile time} risks creating a task that requires more energy than the device can buffer in its energy storage. Such a large task would fail to complete unless the environment provides sufficient incoming energy to supplement the energy in the capacitor. To eliminate this risk, the programmer (or the compiler) must decompose the program conservatively into many small tasks, at the cost of \emph{higher task transition overhead}.

One approach to decrease the task count, without assuming a minimum level of
incoming power always available, is to predict task energy consumption and
choose a decomposition specific to the device capacitor size~\cite{cleancut}. However, statically predicting energy of arbitrary input-dependent code with peripheral access is a problem without a general solution. Furthermore, a static decomposition approach prevents portability across devices with different storage capacitors.

@@

Task coalescing removes the burden from the programmer, because it accepts any task decomposition and improves it dynamically. To further reduce the programmer effort, we propose a compiler pass for \emph{automatic decomposition} of programs into (small) \emph{atomic} tasks. The compiler identifies non-volatile variables shared across tasks as tasks are created, and instruments reads and writes of those variables, using memory virtualization, to keep the data consistent in the presence of power loss. \textcolor{red}{Despite being limited to a subset of the C language, the automatic task decomposition allowed us to port several	applications to an intermittent platform with a moderate effort.}

@@

\begin{itemize}
	\item \textcolor{red}{XXX} adaptive coalescing strategies for selecting boundaries to coalesce based on past execution behaviour;
	\item Software memory virtualisation for transiently-powered devices that preserves atomicity of coalesced tasks;
	\item Task splitting startegy that preserve forward progress and memory consistency;
\end{itemize}
